---
title: "Simulations for MMPD"
author: "Sayani Gupta"
date: "15/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
library(ggridges)
library(viridis)
library(tsibble)
library(gravitas)
library(ggpubr)
library(readr)

opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = "figure/", fig.align = "center", fig.show = "hold",
  cache = TRUE, cache.path = "cache/",
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
knitr::opts_knit$set(root.dir = here::here())
read_chunk('main.R')
```

```{r external, include = FALSE}

```

```{r load}

```

```{r rank_harmony}

```

```{r global_threshold}

```


# Simulation from a Normal distribution
all combinations ~ N(5, 10)

# Simulations from another Normal distribution
all combinations ~ N(800, 40)

# Simulations from a Cauchy distribution

# Simulations from a Exponential distribution

If these look correct

Then check for different levels

same four things

Think about if they would have a range always

check MMPD from start!!!!!!!!!!

max of standard gumbel distribution







